Symmetric Circuits for Rank Gates

### 2018-04-06 sections/transparency-and-unique-extensions.tex
Another general edit. I have changed isomorphism-equivilent to isomorphic. I have also changed the "symmetric circuit" section in order to justify this choice of notation.

\documentclass[../paper.tex]{subfiles}
\begin{document}
Recall that each gate $g$ in a cricuit  is associated with a structure
$D$ with the elements of the relations in $D$, i.e.\ $\ind(\Sigma(g))$
labelled by the children of $g$.  Many of the properties of circuits
that are central to our construction, and we wish to be able to decide
efficiently, particularly those that reference circuit automorphisms or
syntactic-equivalence, depend on the existence of isomorphisms that agrees with a
given permutation of the gates labelling the structure.  This introduces a
potential difficulty as there is no known polynomial-time algorithm for the
general isomorphism problem, but we we require many of these circuit properties
to be polynomial-time decidable in order to define our translation from circuits
to formulas (see Section~\ref{sec:circuits-to-formulas}).  This is not a problem
if we restrict our attention to circuits with symmetric gates, as
in~\cite{AndersonD17}, as in this case the structures of interest are
sets, and set equivalence is clearly decidable in polynomial-time.  When we
allow for circuits to be defined over bases including non-symmetric functions it
is not obvious that these properties can be decided in polynomial-time.

@@

  We now define a function $f: \ind(g) \rightarrow \ind(h)$ as follows. Let
  $\{x_1, \ldots, x_{\vert \ind(g)\vert}\}$ be an ordering of $\ind(g)$. For
  each $i \in \vert \ind(g) \vert$ select $y_i \in \ind(h)$ such that $y_i \neq
  y_j$ for all $j \in [i-1]$ and $L(g)(x_i) \equiv L(h)(y_i)$. If at any point
  we cannot select an appropriate $y_i$, halt and output that $g \not\equiv h$.
  For each $i \in \vert \ind(g) \vert$ let $f(x_i) = y_i$. From the definition
  of the function $f$ is injective. Check if $f$ is surjective. If not, halt and
  output that $g \not\equiv h$. We thus have that $f$ is a bijective function
  and for all $x \in \ind(g)$, $L(g)(x) \equiv L(h)(f(x))$. Check if $g$
  and $h$ are both symmetric gates. If so, $f$ witnesses the required
  isomorphism-equivalence and so we halt and output that $g \equiv h$.

@@

  We now show that $C_\equiv$ is reduced. Suppose $[g], [h] \in G_\equiv$ and
  suppose $[g] \equiv [h]$. If $[g]$ and $[h]$ are both input or output gates
  then $[g] = [h]$. Suppose $[g]$ and $[h]$ are internal gates that are not
  output gates. Then $\Sigma (g) = \Sigma_\equiv ([g]) = \Sigma_\equiv([h]) =
  \Sigma(h)$. There exists $g', h' \in G$ such that $g' \equiv g$ and $h' \equiv
  h$, and $L_\equiv ([g]) = L(g') /_\equiv$ and $L_\equiv ([h]) = L(h')
  /_\equiv$. It follows from $[g] \equiv [h]$ that $L(g') /_\equiv$ is
  isomorphism-equivalent to $L(h') /_\equiv$. It follows that $g' \equiv h'$,
  and so $[g] = [h]$.

  Let $\sigma \in \sym_n$ and suppose there exists $\pi \in \aut(C)$ extending
  $\sigma$. Let $\pi_\equiv = \pi /_\equiv$. We now show that $\pi_\equiv$ is an
  automorphism of $C_\equiv$ extending $\sigma$. It is easy to see that
  $\pi_\equiv$ is a bijection from $G_\equiv$ to $G_\equiv$ that preserves
  syntactic-equivalence, and is thus is a well-defined function. Let $[g] \in
  G_\equiv$. We have that $\Sigma_\equiv (\pi_\equiv [g]) = \Sigma_\equiv ([\pi
  g]) = \Sigma (\pi g) = \Sigma (g) = \Sigma_\equiv ([g])$. It is easy to check
  the automorphism conditions for input gates. Suppose $[g]$ is an internal
  gate. Let $g' \in [g]$ be such that $L_\equiv([g]) = L(g') /_\equiv$ and $h
  \in [g]$ be such that $L_\equiv(\pi_\equiv[g]) = L_\equiv ([\pi h]) = L(\pi h)
  /_\equiv$. We have that $\pi L(g')$ is isomorphism-equivalent to $L(\pi g')$,
  and it follows that $(\pi L(g'))/_\equiv$ is isomorphism-equivalent to $L(\pi
  g') /_\equiv$. We then have $(\pi L(g'))/_\equiv = \pi_\equiv (L(g') /_\equiv)
  = \pi_\equiv L_\equiv ([g])$ and, since $g' \equiv h$ and so $\pi g' \equiv
  \pi h$, we have that $L(\pi g')/_\equiv$ is isomorphism-equivalent to $L(\pi
  h) /_\equiv = L_\equiv(\pi_\equiv[g])$. It follows that $\pi_\equiv L_\equiv
  ([g])$ is isomorphism-equivalent to $L_\equiv(\pi_\equiv [g]) /_\equiv$.
  Suppose $[g]$ is an output gate. Then for $\vec{a} \in \dom (\Omega)$,
  $\pi_\equiv \Omega_\equiv (\vec{a}) = [\pi \Omega (\vec{a})] = [\Omega (\sigma
  \vec{a})] = \Omega_\equiv (\sigma \vec{a})$. It follows that if $C$ is
  symmetric then $C_\equiv$ is symmetric.

@@

  Let $h$ be an internal gate in the circuit and assume we have defined
  $\pi'(g)$ for every gate $g$ of depth less than $h$. Let $h'$ be a gate in the
  circuit such that $\Sigma_\equiv(h) = \Sigma_\equiv (h')$, $\pi' L(h)$ is
  isomorphism-equivalent to $L_\equiv(h')$ and, if $h$ is an output gate then
  $h'$ is an output gate such that $\sigma \Omega^{-1}_\equiv(h) =
  \Omega^{-1}_\equiv(h')$. Suppose there is another gate $h''$ in the circuit
  that meets all of those criteria as well. But then it follows that $h'' \equiv
  h'$. Since $C_\equiv$ is reduced, we then have that $h'' = h'$.  We thus have
  that the choice of $h'$, if it exists, is unique. Note that, since $C$ has
  unique labels $C_\equiv$ has unique labels, and so unique extensions, and so
  we have that $\pi' L_\equiv(h)$ is isomorphism-equivalent to $L_\equiv(h')$
  if, and only if, $L_\equiv(h')^{-1}\pi'L_\equiv(h)$ acts on $\ind(h)$ like an
  automorphism of $\str{h}$. This is easy to determine. If no such $h'$ exists,
  halt and output that there is no automorphism extending $\sigma$. Otherwise,
  let $\pi'(h) = h'$.

@@

  We note that for $i \in [2]$, $L(g^i_{\rank})(p,q)= g^i_\land$ if, and only
  if, $(p,q)$ is an edge in $B_i$ and $L(g^i_{\rank})(p,q) g^i_\lor$ if, and
  only if, $(p,q)$ is not an edge in $B_i$. Let $\pi_{(1,2)} : G \rightarrow G$
  be the function that fixes the output gate, and for each symbol $s$ swaps the
  gate $g^1_s$ with the gate $g^2_s$. It is easy to see that $\pi_{(1,2)}$ is a
  bijection. Moreover, $C$ is symmetric if, and only if, $\pi_{(1,2)}$ is an
  automorphism of the circuit extending the transposition $(1,2)$, if and only
  if, $(1,2) (L(g^1_{\rank}))$ is isomorphism-equivalent to $L(g^2_{\rank})$ if,
  and only if, $B_1$ and $B_2$ are bipartite-isomorphic.

>>>>

\documentclass[../paper.tex]{subfiles}
\begin{document}
Recall that each gate $g$ in a cricuit is associated with a structure $D$ with
the elements of the relations in $D$, i.e.\ $\ind(\Sigma(g))$ labelled by the
children of $g$. Many of the properties of circuits that are central to our
construction, and we wish to be able to decide efficiently, particularly those
that reference circuit automorphisms or syntactic-equivalence, depend on the
existence of isomorphism of structures that agrees with a given permutation of
the gates labelling the structure. This introduces a potential difficulty as
there is no known polynomial-time algorithm for the general isomorphism problem,
but we we require many of these circuit properties to be polynomial-time
decidable in order to define our translation from circuits to formulas (see
Section~\ref{sec:circuits-to-formulas}). This is not a problem if we restrict
our attention to circuits with symmetric gates, as in~\cite{AndersonD17}, as in
this case the structures of interest are sets, and set equivalence is clearly
decidable in polynomial-time. When we allow for circuits to be defined over
bases including non-symmetric functions it is not obvious that these properties
can be decided in polynomial-time.

@@

  We now define a function $f: \ind(g) \rightarrow \ind(h)$ as follows. Let
  $\{x_1, \ldots, x_{\vert \ind(g)\vert}\}$ be an ordering of $\ind(g)$. For
  each $i \in \vert \ind(g) \vert$ select $y_i \in \ind(h)$ such that $y_i \neq
  y_j$ for all $j \in [i-1]$ and $L(g)(x_i) \equiv L(h)(y_i)$. If at any point
  we cannot select an appropriate $y_i$, halt and output that $g \not\equiv h$.
  For each $i \in \vert \ind(g) \vert$ let $f(x_i) = y_i$. From the definition
  of the function $f$ is injective. Check if $f$ is surjective. If not, halt and
  output that $g \not\equiv h$. We thus have that $f$ is a bijective function
  and for all $x \in \ind(g)$, $L(g)(x) \equiv L(h)(f(x))$. Check if $g$ and $h$
  are both symmetric gates. If so, $f$ is an isomorphism from $L(g)$ to $L(h)$,
  and so we halt and output that $g \equiv h$.

@@

  We now show that $C_\equiv$ is reduced. Suppose $[g], [h] \in G_\equiv$ and
  suppose $[g] \equiv [h]$. If $[g]$ and $[h]$ are both input or output gates
  then $[g] = [h]$. Suppose $[g]$ and $[h]$ are internal gates that are not
  output gates. Then $\Sigma (g) = \Sigma_\equiv ([g]) = \Sigma_\equiv([h]) =
  \Sigma(h)$. There exists $g', h' \in G$ such that $g' \equiv g$ and $h' \equiv
  h$, and $L_\equiv ([g]) = L(g') /_\equiv$ and $L_\equiv ([h]) = L(h')
  /_\equiv$. It follows from $[g] \equiv [h]$ that $L(g') /_\equiv$ is
  isomorphic to $L(h') /_\equiv$. From this it follows that $g' \equiv h'$, and
  so $[g] = [h]$.

  Let $\sigma \in \sym_n$ and suppose there exists $\pi \in \aut(C)$ extending
  $\sigma$. Let $\pi_\equiv = \pi /_\equiv$. We now show that $\pi_\equiv$ is an
  automorphism of $C_\equiv$ extending $\sigma$. It is easy to see that
  $\pi_\equiv$ is a bijection from $G_\equiv$ to $G_\equiv$ that preserves
  syntactic-equivalence, and is thus is a well-defined function. Let $[g] \in
  G_\equiv$. We have that $\Sigma_\equiv (\pi_\equiv [g]) = \Sigma_\equiv ([\pi
  g]) = \Sigma (\pi g) = \Sigma (g) = \Sigma_\equiv ([g])$. It is easy to check
  the automorphism conditions for input gates. Suppose $[g]$ is an internal
  gate. Let $g' \in [g]$ be such that $L_\equiv([g]) = L(g') /_\equiv$ and $h
  \in [g]$ be such that $L_\equiv(\pi_\equiv[g]) = L_\equiv ([\pi h]) = L(\pi h)
  /_\equiv$. We have that $\pi L(g')$ is isomorphic to $L(\pi g')$, and it
  follows that $(\pi L(g'))/_\equiv$ is isomorphic to $L(\pi g') /_\equiv$. We
  then have $(\pi L(g'))/_\equiv = \pi_\equiv (L(g') /_\equiv) = \pi_\equiv
  L_\equiv ([g])$ and, since $g' \equiv h$ and so $\pi g' \equiv \pi h$, we have
  that $L(\pi g')/_\equiv$ is isomorphic to $L(\pi h) /_\equiv =
  L_\equiv(\pi_\equiv[g])$. It follows that $\pi_\equiv L_\equiv ([g])$ is
  isomorphic to $L_\equiv(\pi_\equiv [g]) /_\equiv$. Suppose $[g]$ is an output
  gate. Then for $\vec{a} \in \dom (\Omega)$, $\pi_\equiv \Omega_\equiv
  (\vec{a}) = [\pi \Omega (\vec{a})] = [\Omega (\sigma \vec{a})] = \Omega_\equiv
  (\sigma \vec{a})$. It follows that if $C$ is symmetric then $C_\equiv$ is
  symmetric.

@@

  Let $h$ be an internal gate in the circuit and assume we have defined
  $\pi'(g)$ for every gate $g$ of depth less than $h$. Let $h'$ be a gate in the
  circuit such that $\Sigma_\equiv(h) = \Sigma_\equiv (h')$, $\pi' L(h)$ is
  isomorphic to $L_\equiv(h')$ and, if $h$ is an output gate then $h'$ is an
  output gate such that $\sigma \Omega^{-1}_\equiv(h) = \Omega^{-1}_\equiv(h')$.
  Suppose there is another gate $h''$ in the circuit that meets all of those
  criteria as well. But then it follows that $h'' \equiv h'$. Since $C_\equiv$
  is reduced, we then have that $h'' = h'$. We thus have that the choice of
  $h'$, if it exists, is unique. Note that, since $C$ has unique labels
  $C_\equiv$ has unique labels, and so unique extensions, and so we have that
  $\pi' L_\equiv(h)$ is isomorphic to $L_\equiv(h')$ if, and only if,
  $L_\equiv(h')^{-1}\pi'L_\equiv(h)$ acts on $\ind(h)$ like an automorphism of
  $\str{h}$. This is easy to determine. If no such $h'$ exists, halt and output
  that there is no automorphism extending $\sigma$. Otherwise, let $\pi'(h) =
  h'$.

@@

  We note that for $i \in [2]$, $L(g^i_{\rank})(p,q)= g^i_\land$ if, and only
  if, $(p,q)$ is an edge in $B_i$ and $L(g^i_{\rank})(p,q) g^i_\lor$ if, and
  only if, $(p,q)$ is not an edge in $B_i$. Let $\pi_{(1,2)} : G \rightarrow G$
  be the function that fixes the output gate, and for each symbol $s$ swaps the
  gate $g^1_s$ with the gate $g^2_s$. It is easy to see that $\pi_{(1,2)}$ is a
  bijection. Moreover, $C$ is symmetric if, and only if, $\pi_{(1,2)}$ is an
  automorphism of the circuit extending the transposition $(1,2)$, if and only
  if, $(1,2) (L(g^1_{\rank}))$ is isomorphic to $L(g^2_{\rank})$ if, and only
  if, $B_1$ and $B_2$ are bipartite-isomorphic.
